A lottery model of density-dependent selection

### 2017-11-08 manuscript.tex
l

A variety of different measures of fitness are used to quantify the effects of selection. Some widely used examples in evolutionary ecology are expected lifetime reproductive ratio $R_0$,  intrinsic growth rate $r$, saturation population density (often labeled ``$K$'') \citep{benton_2000}, and invasion fitness \citep{metz_1992}. In addition, ``relative fitness'' is the standard in much of evolutionary biology, particularly evolutionary genetics, where attention is generally restricted to relative genotypic proportions \cite[pp. 468]{barton_2007}.

In principle these fitness measures should all be grounded in population ecology, because ultimately selection operates through the environmentally-dependent differential reproduction and survival of individuals \citep{metcalf_2007}. Widely-used ``absolute'' measures of fitness like $r$ and $K$ are derived directly from population ecology. By contrast, relative fitness is almost entirely divorced from population ecology. This discrepancy runs deeper than a simple difference in emphasis on absolute densities rather than relative frequencies.

In uncrowded populations there is no discrepancy. Relative fitness simply represents  differences in $r$ ($r$-selection) \citep[pp. 26]{crow_1970}. The discrepancy emerges in crowded populations. Then the relative fitness description of selection assumes that $N$ is fixed, and fitness strictly involves the ability to increase in frequency at the expense of others (e.g. in the Wright-Fisher model). On the other hand, from quite general population ecological considerations, MacArthur famously argued that ``fitness is $K$'' in crowded populations ($K$-selection) \citep[pp. 149]{macarthur_1967}. To appreciate the import of MacArthur's argument, and see the nature of the discrepancy between these views of selection, we now briefly review MacArthur's argument.

>>>>

There are a variety of different measures of fitness. Some widely used examples in evolutionary ecology are expected lifetime reproductive ratio $R_0$, intrinsic growth rate $r$, saturation population density (often labeled ``$K$'') \citep{benton_2000}, and invasion fitness \citep{metz_1992}. In addition, ``relative fitness'' is the standard in much of evolutionary biology, particularly evolutionary genetics, where attention is generally restricted to relative genotypic proportions \cite[pp. 468]{barton_2007}.

Despite this variety, fitness measures are different ways to environmentally-dependent quantify the differential reproduction and survival of individuals \citep{metcalf_2007}.

relative fitness is almost entirely divorced from population ecology.

In uncrowded populations, there is no discrepancy. Relative fitness simply represents  differences in $r$ ($r$-selection) \citep[pp. 26]{crow_1970}.

Absolute measures of fitness like $r$ and $K$ are derived directly from population ecology. By contrast,

This discrepancy runs deeper than a simple difference in emphasis on absolute densities rather than relative frequencies.

 The discrepancy emerges in crowded populations. Then the relative fitness description of selection assumes that $N$ is fixed, and fitness strictly involves the ability to increase in frequency at the expense of others (e.g. in the Wright-Fisher model). On the other hand, from quite general population ecological considerations, MacArthur famously argued that ``fitness is $K$'' in crowded populations ($K$-selection) \citep[pp. 149]{macarthur_1967}. To appreciate the import of MacArthur's argument, and see the nature of the discrepancy between these views of selection, we now briefly review MacArthur's argument.
